Dominance Regions for Rank Two Cluster Algebras

### 2022-05-08 dominance_regions.tex
More edits?
Citing Qin

\documentclass[a3paper]{amsart}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{color}
%\usepackage{ulem}
\usepackage[margin=1in]{geometry}
\usepackage{tikz}
\usetikzlibrary{math}
\usepackage[bookmarks=true, bookmarksopen=true, bookmarksdepth=3,bookmarksopenlevel=2, colorlinks=true, linkcolor=blue, citecolor=blue, filecolor=blue, menucolor=blue, urlcolor=blue]{hyperref}

@@

  Important examples of pointed elements are the \emph{cluster monomials} of $\cA(b,c)$, i.e. the elements that can be written in the form $x_k^{\alpha_k}x_{k+1}^{\alpha_{k+1}}$ for some $k\in\ZZ$ and $\alpha_k,\alpha_{k+1}$ non-negative integers.
  Indeed, by \cite[Lemma 3.4.12]{qin} cluster monomials are part of any pointed basis in any cluster algebra.
  We achieve the same conclusion in our setting by elementary calculations (c.f. Lemma~\ref{le:cluster monomials}).

  Elements of any pointed basis are parametrized by their $\bfg$-vectors.
  Qin introduced a partial order on $\bfg$-vectors called the \emph{dominance order} and showed that given two pointed bases $x=\{x_\lambda\}$, $y=\{y_\lambda\}$, $y_\mu$ can appear in the expansion of $x_\lambda$ in terms of $y$ only if $\lambda$ dominates $\mu$ (cf. \cite[Theorem 1.2.1]{qin}).

@@

The paper is organized as follows.
In Section~\ref{sec:chebyshev}, we collect useful results related to two-parameter Chebyshev polynomials which support our main calculations.
Section~\ref{sec:tropical} contains calculations related to the transformation of $\bfg$-vectors under mutations.
Section~\ref{sec:dominance inequalities} proves Theorem~\ref{th:dominance inequalities}.
Section~\ref{sec:dominance vertices} proves Corollary~\ref{cor:dominance vertices}.
Section~\ref{sec:maximum support} proves Theorem~\ref{th:maximum support}.
The paper ends with Section~\ref{sec:affine} interpreting the dominance polygon in terms of generalized minors in the cases where $bc=4$.

@@

  \begin{theorem}
    \label{th:dominance}
    \cite{qin}
    \sayDR{Quote Fan's result here.}
  \end{theorem}

@@

  For $\mu_1-\lambda_1\ge 0$, the inequalities \eqref{ieq1} become more restrictive as $k$ gets larger since the sequence $\frac{-u_{k+1}^-}{u_k^+}$ of negative slopes is monotonically increasing (c.f. Lemma~\ref{le:limits}).
  Thus, following Lemma~\ref{le:limits}, in the limit as $k\to\infty$, we obtain the inequality 
  \[ 0 \le \frac{-bc-\sqrt{bc(bc-4)}}{2b}(\mu_0-\lambda_0)-(\mu_1-\lambda_1) \]
  or, equivalently,
  \[ 0 \le -(\mu_0-\lambda_0)-\frac{bc-\sqrt{bc(bc-4)}}{2c}(\mu_1-\lambda_1) \]
  determining a boundary of $\cP_\lambda$.

@@

  Finally, the inequalities \eqref{ieq3} become more restrictive for $\mu_1\le0$ as $k$ gets larger since the sequence $\frac{-u_k^-}{u_{k-1}^+}$ of negative slopes is monotonically increasing (c.f. Lemma~\ref{le:limits}) and the intersection with \eqref{ieq2} (for $k+1$) moves to the right on the $\mu_0$-axis as $k$ increases.
  Thus, following Lemma~\ref{le:limits}, in the limit as $k\to\infty$, we obtain the inequality
  \[ -(\mu_0-\lambda_0)-\frac{bc-\sqrt{bc(bc-4)}}{2c}(\mu_1-\lambda_1) \le \frac{bc+\sqrt{bc(bc-4)}}{2c}c\lambda_0+b\lambda_1 \]
  determining a boundary of $\cP_\lambda$.

@@

  The following result is proven in \cite[Lemma 2.4]{rupel} with appropriate notation changes, c.f. Theorem~\ref{th:dominant Dyck path recursion} for a slight modification and explicit proof.
  \begin{lemma}\mbox{}
    \label{le:maximal Dyck path recursion}
    \begin{enumerate}
      \item If $a_2 < c a_1$, then $\mu_{1,c}\big(D^{max}[a_1,a_2]\big)=D^{max}[ca_1-a_2,a_1]$.
      \item If $a_1 < b a_2$, then $\mu_{2,b}\big(D^{max}[a_1,a_2]\big)=D^{max}[a_2,ba_2-a_1]$.
    \end{enumerate}
  \end{lemma}

>>>>

\documentclass{amsart}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{color}
%\usepackage{ulem}
\usepackage[margin=1in]{geometry}
\usepackage{tikz}
\usetikzlibrary{math}
\usepackage[bookmarks=true, bookmarksopen=true, bookmarksdepth=3,bookmarksopenlevel=2, colorlinks=true, linkcolor=blue, citecolor=blue, filecolor=blue, menucolor=blue, urlcolor=blue]{hyperref}

@@

  Important examples of pointed elements are the \emph{cluster monomials} of $\cA(b,c)$, i.e. the elements that can be written in the form $x_k^{\alpha_k}x_{k+1}^{\alpha_{k+1}}$ for some $k\in\ZZ$ and $\alpha_k,\alpha_{k+1}$ non-negative integers.
  Indeed, by \cite[Lemma 3.4.12]{qin} cluster monomials are part of any pointed basis in any cluster algebra.
  We achieve the same conclusion in our setting by elementary calculations (cf. Lemma~\ref{le:cluster monomials}).

  Elements of any pointed basis are parametrized by their $\bfg$-vectors.
  Qin introduced a partial order $\preceq$ on $\bfg$-vectors called the \emph{dominance order} and showed the following result characterizing pointed bases which motivates this paper.

  \begin{theorem}
    \label{th:dominance}
    \cite[Theorem 1.2.1]{qin}
    Let $x=\{x_\lambda\}$ and $y=\{y_\lambda\}$ be pointed bases of $\cA(b,c)$.
    Then for each $\lambda\in\ZZ^2$, there exist scalars $a_{\lambda,\mu}\in\ZZ$ for $\mu\prec\lambda$ such that
    \[y_\lambda=x_\lambda+\sum_{\mu\prec\lambda} a_{\lambda,\mu} x_\mu.\]
  \end{theorem}

@@

The paper is organized as follows.
In Section~\ref{sec:chebyshev}, we collect useful results related to two-parameter Chebyshev polynomials which support our main calculations.
Section~\ref{sec:tropical} contains calculations related to the transformation of $\bfg$-vectors under mutations.
Section~\ref{sec:dominance inequalities} proves Theorem~\ref{th:dominance inequalities}.
Section~\ref{sec:dominance vertices} proves Corollary~\ref{cor:dominance vertices}.
Section~\ref{sec:maximum support} proves Theorem~\ref{th:maximum support}.
The paper ends with Section~\ref{sec:affine} interpreting the dominance polygon in terms of generalized minors in the cases where $b=c=2$.

@@

@@

  For $\mu_1-\lambda_1\ge 0$, the inequalities \eqref{ieq1} become more restrictive as $k$ gets larger since the sequence $\frac{-u_{k+1}^-}{u_k^+}$ of negative slopes is monotonically increasing (cf. Lemma~\ref{le:limits}).
  Thus, following Lemma~\ref{le:limits}, in the limit as $k\to\infty$, we obtain the inequality 
  \[ 0 \le \frac{-bc-\sqrt{bc(bc-4)}}{2b}(\mu_0-\lambda_0)-(\mu_1-\lambda_1) \]
  or, equivalently,
  \[ 0 \le -(\mu_0-\lambda_0)-\frac{bc-\sqrt{bc(bc-4)}}{2c}(\mu_1-\lambda_1) \]
  determining a boundary of $\cP_\lambda$.

@@

  Finally, the inequalities \eqref{ieq3} become more restrictive for $\mu_1\le0$ as $k$ gets larger since the sequence $\frac{-u_k^-}{u_{k-1}^+}$ of negative slopes is monotonically increasing (cf. Lemma~\ref{le:limits}) and the intersection with \eqref{ieq2} (for $k+1$) moves to the right on the $\mu_0$-axis as $k$ increases.
  Thus, following Lemma~\ref{le:limits}, in the limit as $k\to\infty$, we obtain the inequality
  \[ -(\mu_0-\lambda_0)-\frac{bc-\sqrt{bc(bc-4)}}{2c}(\mu_1-\lambda_1) \le \frac{bc+\sqrt{bc(bc-4)}}{2c}c\lambda_0+b\lambda_1 \]
  determining a boundary of $\cP_\lambda$.

@@

  The following result is proven in \cite[Lemma 2.4]{rupel} with appropriate notation changes, cf. Theorem~\ref{th:dominant Dyck path recursion} for a slight modification and explicit proof.
  \begin{lemma}\mbox{}
    \label{le:maximal Dyck path recursion}
    \begin{enumerate}
      \item If $a_2 < c a_1$, then $\mu_{1,c}\big(D^{max}[a_1,a_2]\big)=D^{max}[ca_1-a_2,a_1]$.
      \item If $a_1 < b a_2$, then $\mu_{2,b}\big(D^{max}[a_1,a_2]\big)=D^{max}[a_2,ba_2-a_1]$.
    \end{enumerate}
  \end{lemma}
